---
title: "Project2"
author: "Rashmi Kadam, Dionte Watie"
date: "7/6/2021"
output:
  pdf_document:
    toc: true
  github_document:
    toc: true
  params: 
    day: day
    weekday: weekday
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(fig.path=paste0(day, "_files/figure-gfm/"))

```


# Introduction

For this study we will be aiming to predict the number of bike users. The bike users have been split into two groups that will be the target variables (response), casual bikers that rent bikes casually and registered bikers that rent bikes regularly. The predictor variables that will be in question are:

* weekday (day of the week)
* season
* yr (year)
* holiday (whether it is a holiday or not)
* weathersit (weather: rainy, snowy, clear, cloudy)
* mnth
* atemp 
* windspeed


The response and predictor variables will be used in various Multiple Linear Regression Models, Logistics Models, and Tree fits. The models will then be tested against the testing data set and the results will determine which model would be best to use for prediction.


```{r}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(ggpubr)
library(caret)
library(randomForest)
```



```{r}
set.seed(1)

# read Bike data
bikeData <- read_csv("day.csv")

# filtering weekday data
bikeDataWD <- bikeData %>% filter(weekday == 1)

```
```{r}

Correlation <- cor(select(bikeDataWD, casual, registered, cnt,holiday, mnth, season,  weathersit ,yr, temp, atemp, hum, windspeed))

corrplot(Correlation)

```

```{r}
#Running the model

bikeDataM <- bikeDataWD %>% select (season, holiday, mnth, weathersit, atemp, windspeed, casual, registered, cnt, yr)

#
bikeDataM$mnth <- as.factor(bikeDataM$mnth)

bikeDataM$season <- factor(bikeDataM$season,
                          levels = c("1", "2","3","4") , 
                          labels = c("Spring","Summer","Fall","Winter"))

bikeDataM$holiday <- factor(bikeDataM$holiday, 
                           levels = c("0", "1") , 
                           labels = c("Working Day","Holiday"))


bikeDataM$weathersit <- factor(bikeDataM$weathersit,
                          levels = c("1", "2","3","4") , 
                          labels = c("Good:Clear/Sunny","Moderate:Cloudy/Mist","Bad: Rain/Snow/Fog",
                                     "Worse: Heavy Rain/Snow/Fog"))

bikeDataM$yr <- factor(bikeDataM$yr,
                      levels = c("0", "1") , 
                      labels = c("2011","2012"))

```

```{r}

train <- sample(1:nrow(bikeDataM), size = nrow(bikeDataM)*0.7)
test <- dplyr::setdiff(1:nrow(bikeDataM), train)
bikeDataTrain <- bikeDataM[train, ]
bikeDataTest <- bikeDataM[test, ]


summary(bikeDataTrain)

```
 
```{r}
#Rashmi- side by side bar plots for year, month and count
  
Year <- bikeDataTrain$yr

ggplot(bikeDataTrain, aes(fill=Year, y=cnt, x=mnth)) + 
    geom_bar(position="dodge", stat="identity") + xlab("Months") + ylab('Total Users')

```


```{r}
#Rashmi-boxplot of season versus count

ggplot(bikeDataTrain, aes(x = season,y=cnt)) +
  geom_boxplot(fill="steelblue") +ylab('Total Users')


```


```{r}
#Rashmi-- scatter plots for casual and registered versus actual temperature

mintemp <- -16
maxtemp <- 50

bikeDataTrain$atemp = bikeDataTrain$atemp * (maxtemp - mintemp) + mintemp
bikeDataTrain$atemp

  cTemp <- ggplot(bikeDataTrain,aes(x=atemp, y=casual)) + geom_point() + geom_smooth() + ylim(0, 7000) 
  rTemp <- ggplot(bikeDataTrain, aes(x=atemp, y=registered)) + geom_point() + geom_smooth() + ylim(0, 7000) 
  
  ggarrange(cTemp, rTemp, labels = c("Casual Users", "Registered Users"), ncol = 2, nrow = 1)



```


```{r}
#Dionte -- bar plots split by casual and registered users for season and holiday
rSeason <- ggplot(bikeDataTrain, aes(fill = holiday, x = season,y = registered,)) + geom_bar(position= 'dodge',stat = 'identity')

cSeason <- ggplot(bikeDataTrain, aes(fill = holiday, x = season,y = casual,)) + geom_bar(position= 'dodge',stat = 'identity')

ggarrange(cSeason, rSeason, labels= c("Casual Users", "Registered Users"), ncol = 2, nrow = 1)

```


```{r}
#Dionte -- Density plot for weathersit by year 
weather <- ggplot(bikeDataTrain, aes(x= weathersit))
weather + geom_density(adjust= 0.5, alpha= 0.5, aes(fill= Year), kernel="gaussian")

```

```{r}
#Dionte -- ECDF plot for count by year
cntPlot <- ggplot(bikeDataTrain, aes(x= cnt))
cntPlot + stat_ecdf(geom = 'step', aes(color= Year)) + ylab("ECDF")


``` 


```{r}
#Summary Statistics by Dionte
#variance, stdev, mean, and median of casual users by month
statsCasual <- bikeDataTrain %>% 
  group_by(mnth) %>% 
  summarise(avg = mean(casual), 
            med = median(casual), 
            var = var(casual), 
            stDev = sd(casual))
statsCasual
#variance, stdev, mean, and median of registered users by month
statsRegistered <- bikeDataTrain %>% 
  group_by(mnth) %>% 
  summarise(avg = mean(registered), 
            med = median(registered), 
            var = var(registered), 
            stDev = sd(registered))
statsRegistered
#variance, stdev, mean, and median of total bike users by season
statsCnt <- bikeDataTrain %>% 
  group_by(season) %>% 
  summarise(avg = mean(cnt), 
            med = median(cnt), 
            var = var(cnt), 
            stDev = sd(cnt))
statsCnt
```


```{r}
#Calculating z statistic
tapply(bikeDataTrain$casual, INDEX = bikeDataTrain$weathersit, FUN = function(x){x -mean(x)/sd(x)})
```



```{r}
# summary statistics by Rashmi
#min max stdev and mean of feeling temperature by season
statsAtemp <- bikeDataTrain %>%
  group_by(season) %>%
  summarise(
    atemp.min = min(atemp),
    atemp.max = max(atemp),
    atemp.med = median(atemp),
    atemp.stdev = sd(atemp),
    atemp.mean = mean(atemp))
statsAtemp

#min max stdev and mean of total bike users per year
statsYear<- bikeDataTrain %>%
  group_by(yr) %>%
  summarise(
    cnt.min = min(cnt),
    cnt.max = max(cnt),
    cnt.med = median(cnt),
    cnt.stdev = sd(cnt),
    cnt.mean = mean(cnt)) 
statsYear

#min max stdev and mean of total bike users per holiday
statsHoliday<- bikeDataTrain %>%
  group_by(holiday) %>%
  summarise(
    cnt.min = min(cnt),
    cnt.max = max(cnt),
    cnt.med = median(cnt),
    cnt.stdev = sd(cnt),
    cnt.mean = mean(cnt)) 
statsHoliday

#contingency table 



```

Fitting models

<<<<<<< HEAD
```{r}
#Fitting multiple regression model by Rashimi
=======
```{r warning=FALSE}
#Fitting multiple regression models
>>>>>>> dc1a8cec6ebd6fc2edad69cee4e14286865120cd

bikeDataTrainF <- bikeDataTrain %>% select(cnt,season,weathersit,atemp,yr)
bikeDataTestF <- bikeDataTest %>% select(cnt,season,weathersit,atemp,yr)

lmRM<-lm(cnt~.,data=bikeDataTrainF)
summary(lmRM)

lmFit <- train(cnt ~ ., data = bikeDataTrainF , 
         method = "lm", 
         preProcess = c("center", "scale"),
         trControl = trainControl(method = "cv", number = 10))

predlm <- predict(lmFit, newdata = dplyr::select(bikeDataTestF,-cnt))

postResample(predlm, bikeDataTestF$cnt)

```

```{r}
#Fitting Multiple Linear Regression model by Dionte
#Using BIC to select predictors for the best fit model
bic_selection = step(
  lm(cnt ~ 1, bikeDataTrain),
  scope = cnt ~ season + holiday + mnth + weathersit + atemp + windspeed + Year,
  direction = "both", k = log(nrow(bikeDataTrain))
)
#Best fit linear regression model
bestLinFit <- lm(cnt ~ atemp + Year + season + weathersit, data = bikeDataTrain)
bestLinFit
summary(bestLinFit)
plot(bestLinFit)
```



Fitting Random forest model

```{r}

set.seed(1)


rfFit <- train(cnt ~ ., 
               method = "rf",
               trControl = trainControl(method = "repeatedcv",
                                        repeats = 3,
                                        number = 10),
               tuneGrid = data.frame(mtry = 1:9),
               data = bikeDataTrainF)
rfFit

rfPred <- predict(rfFit,  newdata = dplyr::select(bikeDataTestF,-cnt))
postResample(rfPred, bikeDataTestF$cnt)


```

Fitted Boosted Tree Model

```{r}
trCtrl <- trainControl(method = "repeatedcv", number = 10, repeats =3)
set.seed(1)
bikeBoostedFit <- train(cnt ~., data = bikeDataTrain,
                     method = "gbm",
                     preProcess = c("center", "scale"),
                     trControl = trCtrl)
bikeBoostedFit

bikeBoostedFitPred <- predict(bikeBoostedFit, newdata = dplyr::select(bikeDataTest, -cnt))
bikeBoostedFitPred
postResample(bikeBoostedFitPred, bikeDataTest$cnt)


```

